---
title: "Usign the fars package"
author: "Ryan Honomichl"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set_global_options, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# overview  
This vignette is written for the package `fars` as part of week 4 assignment of the coursera course "R packages". The aim of this is to describe the functions in the package and their use cases. The packages consists of 5 functions:  

  - `make_filename`  
  
  - `fars_read`  
  
  - `fars_read_years`  
  
  - `fars_summarize_year`  
  
  - `fars_map_state`  

# Data  
The functions provided for you in this assignment will be using data from the US National Highway Traffic Safety Administration's [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars), which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. Once the package is installed 3 example files are accessible as follows:  

```{r access_extdata, eval=FALSE}
system.file('extdata', 'accident_2013.csv.bz2', package = 'fars')
system.file('extdata', 'accident_2014.csv.bz2', package = 'fars')
system.file('extdata', 'accident_2015.csv.bz2', package = 'fars')
```

I will use these files to explain the use of the functions in the package.  

# Functions  
## Using `make_filenam`  

This function is used to construct a file name using the corresponding year. The following is an example to construct a FARS data file name for the year 2013.  

```{r load_required_libraries}
library(fars)
library(maps)
```

```{r use_make_filename}
make_filename(2013)
```

## Using `fars_read`  
This function is a wrapper of the `readr::read_csv` and customized to parse a FARS data file when exists otherwise returns an error. The following is an example of using the function to read the file that I constructed its name in the last example, then try read a file that doesn't exist.   

```{r use_fars_read}
# file exist
dat <- fars_read('accident_2013.csv.bz2')
class(dat)
# file does nott exist
fars_read('accident_1900.csv.bz2')
```

## Using `fars_read_years`

This is a custom function that is used to parse a number of FARS data files for the corresponding input `years`. If first construct file names using `make_filename` then reads the files in a `list` using `fars_read`. The function returns a warning for years that do not exist, as in the following example.  

```{r use_fars_read_years}
# all years exist
yrs <- 2013:2015
dat <- fars_read_years(yrs)
class(dat)
identical(length(yrs), length(dat))
# year 2016 doesn't exist; warning returned
yrs <- 2013:2016
dat <- fars_read_years(yrs)
```

## Using `fars_summarize_years`  
This function provides a numerical summary of the total number of accidents per month for each year. It takes as input a `years` and wraps the previous functionality added to the summary. The following example uses the year 2013 as an input, the function will construct a file name and read the FARS data file then prints a summary.  

```{r use_fars_summarize_years}
yrs <- 2013:2015
fars_summarize_years(yrs)
```

## Using `fars_map_state`  
This function plots the coordinates of individual accidents for a particular `year` on a map of a particular state, `state_num`. An error is returned if the `state.num` is invalid and a message is returned when the number of accidents for this state is 0 at this year.  

```{r use_fars_mape_state, fig.align='center', fig.height=7, fig.width=7}
# valid input; plot returned
fars_map_state(50, 2013)
# invalid input; error returned
fars_map_state(500, 2013)
```

© 2019 GitHub, Inc.
